---
title: "notebook"
author: "Luke Vawter"
date: "December 13, 2018"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
    toc_float: yes
---

`r Sys.Date()`

```{r, echo=FALSE, cache=FALSE}
evaluate <- TRUE
cache.me <- FALSE
run.cedr.acquisition <- FALSE
run.itis.acquisition <- FALSE
run.va_odu.acquisition <- FALSE
run.md.acquistion <- FALSE
```

## Load packages
```{r child= 'sections/load_packages.Rmd', eval=TRUE}

```

## Custom functions and variables
```{r child= 'sections/funcs_and_vars.Rmd', eval=TRUE}

```


##CEDR eventdata, phyto reported count, picophytoplankton reported count
```{r child= 'section/cedr_data.Rmd', eval=TRUE}

```

##VA_ODU eventdata, phyto reported count, picophytoplankton reported count
```{r child= 'sections/va_odu_data.Rmd', eval=TRUE}

```

##MD eventdata and, phyto reported count, picophytoplankton reported count

##assign data to dataframes
```{r child= 'sections/assign_df.Rmd', eval=TRUE}

```

## prep event data
```{r child= 'sections/prep_event.Rmd', eval=TRUE}

```

## CEDR water quality data
```{r child= 'sections/wq_data.Rmd', eval=TRUE}

```

## 3 day window
```{r child= 'sections/3_day_window.Rmd', eval=TRUE}

```

## salinity
```{r child= 'sections/ssalinity.Rmd', eval=TRUE}

```
