---
title: "pibi_long"
author: "Luke Vawter"
date: "December 21, 2018"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
    toc_float: yes
---
supress error warnings
```{r setup, cache=FALSE}
knitr::opts_chunk$set(error = TRUE)

```



#### Download Packages
```{r load_packages, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(stringr)
library(leaflet)
library(lubridate)
library(readxl)
library(httr)
library(RCurl)
library(data.table)
library(jsonlite)
library(rprojroot)
library(ritis)
library(parallel)
```

mmir is Zach Smith's custom package
```{r}
#library(mmir)
```


#### Funcs and Variables
these varaible should be changed by the user as needed
```{r}
#this is the minimum date to be included
min_date = "1-1-2013"
#this is the minimum date to be included sdd the variable todays.date intot his variable if you want the most recent data
max_date = "12-31-2016"
```

these variables references values found in the API guide at http://datahub.chesapeakebay.net (or int eh case of parameters, fonud from trial and error)
```{r}
#number for picoplankton in CEDR api
pico_num = "18"

#number for phytoplankton in CEDR api
phyto_num <- "17"

#these are the parameter codes for the CEDR api
chla=21
doc=34 
pheo=74 
salinity=83

#data_source = ""
```

variables and functions that will likely not be changed
```{r}
project.dir <- rprojroot::find_rstudio_root_file()

clean_string <- function(x) {
  x %>% 
    stringr::str_trim() %>% 
    tolower() %>% 
    stringr::str_replace_all("\\s+", " ") %>% 
    stringr::str_replace_all(" ", "_") %>%  
    if_else(. == "", as.character(NA), .)
}

clean_up <- function(x) {
  x %>% 
    rename_all(clean_string) %>% 
    mutate_if(is.character, funs(clean_string))%>% 
    distinct()
}

url.root <- "http://datahub.chesapeakebay.net/api.JSON"
todays.date <- format(Sys.Date(), "%m-%d-%Y")


dir.create(file.path(project.dir, "data/VA_ODU/"),
           recursive = TRUE, showWarnings = FALSE)

dir.create(file.path(project.dir, "data/CEDR/"),
           recursive = TRUE, showWarnings = FALSE)

dir.create(file.path(project.dir, "data/MD_DNR/"),
           recursive = TRUE, showWarnings = FALSE)

dir.create(file.path(project.dir, "data/water_quality"),
           recursive = TRUE, showWarnings = FALSE)

dir.create(file.path(project.dir, "data/itis"),
           recursive = TRUE, showWarnings = FALSE)


station.vec <- file.path(url.root,
                       "LivingResources",
                       "TidalPlankton",
                       "Reported",
                       min_date,
                       # "1-1-2013",
                       max_date, 
                       # "12-31-2016",
                       phyto_num,
                       "Station") %>% 
  fromJSON() %>% 
  pull(unique(MonitoringLocationId))

```

#### CEDR data



this uploads the sample and event data from cedr to a dataframe(phyto.df)
```{r}

phyto.df <- file.path(url.root,
                      "LivingResources",
                      "TidalPlankton",
                      "Reported",
                       min_date,
                       # "1-1-2013",
                       max_date, 
                       # "12-31-2016",
                      phyto_num,
                      "Station",
                      paste(station.vec, collapse = ",")) %>%
  fromJSON() %>% 
  clean_up()




phyto.df %>% 
  mutate(reportingvalue = as.character(reportingvalue)) %>% 
data.table::fwrite(file.path(rprojroot::find_rstudio_root_file(), "data/CEDR", "cedr_phyto_taxa.csv"))





event.df <- file.path(url.root,
                      "LivingResources",
                      "TidalPlankton",
                      "MonitorEvent",
                       min_date,
                       # "1-1-2013",
                       max_date, 
                       # "12-31-2016",
                      phyto_num,
                      "Station",
                      paste(station.vec, collapse = ",")) %>%
  fromJSON() %>% 
  clean_up()





data.table::fwrite(event.df, file.path(rprojroot::find_rstudio_root_file(), "data/CEDR", "cedr_phyto_event.csv"))
```


COMPARE STEP 1
note: both have 824 elements
```{r}
step_1.df <- event.df

data.table::fwrite(step_1.df, file.path(project.dir, "data/test", "step_1.csv"))

s1_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step1.csv")) %>% clean_up() #%>% mutate(sampledate=as.Date(sampledate))
```



pico data(note that the number above "station" is now 18 rather than 17.  The project id for picoplankton is 18, and 17 for phytoplankton)
```{r}
pico.df <- file.path(url.root,
                      "LivingResources",
                      "TidalPlankton",
                      "Reported",
                       min_date,
                       # "1-1-2013",
                       max_date, 
                       # "12-31-2016",
                       pico_num,
                      "Station",
                      paste(station.vec, collapse = ",")) %>%
  fromJSON() %>% 
  clean_up()


pico.df %>% 
  mutate(reportingvalue = as.character(reportingvalue)) %>% #need this?
data.table::fwrite(file.path(project.dir, "data/CEDR/", "cedr_pico_taxa.csv"))
```


Read csv's to dataframes


You can run the code from here if you've already pulled from the api
CEDR
```{r}
phyto.df <- data.table::fread(file.path(project.dir, "data/CEDR/", "cedr_phyto_taxa.csv")) %>% clean_up() %>% mutate(sampledate=as.Date(sampledate))

event.df <- data.table::fread(file.path(project.dir, "data/CEDR/", "cedr_phyto_event.csv")) %>% clean_up() %>% mutate(sampledate=as.Date(sampledate))

pico.df <- data.table::fread(file.path(project.dir, "data/CEDR/", "cedr_pico_taxa.csv")) %>% clean_up() %>% mutate(sampledate=as.Date(sampledate))

```

#### water quality data

download water quality data:
```{r}
wq.df <- file.path(url.root,
                   "WaterQuality",
                   "WaterQuality",

                   min_date,
                   max_date,
                   # format(min(phyto.df$sampledate) - days(3), "%m-%d-%Y"),
                   # format(max(phyto.df$sampledate) + days(3), "%m-%d-%Y"),
                   "6",  #programIds..."6"" or, "2,4,6",
                   "7,23",#"7,16,23,24",  #projectIds
                   "station",
                   paste(station.vec, collapse = ","),
                   paste(
                     chla,doc,pheo,salinity
                     , sep=",")) %>%
                   #"21,34,74,83") %>% #parameter:21=chla, 34=doc, 74=pheo, 83=salinity
  fromJSON() %>% 
  clean_up()

```
"ProjectId":7,"ProjectIdentifier":"MAIN","ProjectName":"Tidal Mainstem Water Quality Monitoring Project"
"ProjectId":16,"ProjectIdentifier":"PART","ProjectName":"Tidal Non-Traditional Partners"
"ProjectId":23,"ProjectIdentifier":"TRIB","ProjectName":"Tidal Tributary Water Quality Monitoring Project"
"ProjectId":24,"ProjectIdentifier":"TSPECIAL","ProjectName":"Special Tidal Water Quality Monitoring Project"

search for mutant ret3.1
```{r}
wq_mutant.df <- wq.df %>%
  filter(eventid == 425680) #%>%
  #filter(is.na(parameter)) 
```

```{r}
missing_other.df <- wq.df %>%
  mutate(sampledate = as.Date(sampledate)) %>%
  #filter(eventid == 394855 %>% 
  filter(totaldepth <= 5 & station %in% c("tf3.3","tf4.2","tf5.5", "ret3.1","ret4.3","ret3.1")) 
  #filter( station %in% c("tf3.3","tf4.2","tf5.5", "ret3.1","ret4.3","ret3.1") &  (is.na(layer)) |!layer %in% c("s","f","m","p", "ap","b", "bp"))
```

this is a check to see if any wq.df sampledate values fall outside of the date range.  None do
```{r}
print(min(wq.df$sampledate))

print(max(wq.df$sampledate))

narrow_date_range.df <-  wq.df %>%
  mutate(sampledate = as.Date(sampledate)) %>%
  filter( sampledate < as.Date("2013-1-1'") |  sampledate > as.Date("2016-12-31"))


```


```{r}
# step_3.df <- wq_w_pdepth.df
# 
# data.table::fwrite(step_3.df, file.path(project.dir, "data/test", "step_3.csv"))

s3_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step3.csv")) %>% clean_up() #%>% mutate(sampledate=as.Date(sampledate))

```


step 3 anti
```{r}
s3_compare_modified.df <- s3_compare.df %>%
  rename(eventid = wq_eventid)


compare_wq_raw.df <- anti_join(wq.df, s3_compare_modified.df, by = c("eventid"))
```


```{r}
station_check <- s3_compare_modified.df %>%
  filter(station %in% c("sbe5"))
```


```{r}
data.table::fwrite(wq.df, file.path(project.dir, "data/water_quality", "cedr_wq.csv"))
```


```{r}
wq.df <- data.table::fread(file.path(project.dir, "data/water_quality/cedr_wq.csv"),
                            data.table = FALSE,
                           na.strings = c(""))
```

run from war wq(right after CEDR pull) note: this is actually STEP 3
we reduce the wq data to match teh phyto data
```{r}
#create concatinated id for phyto events
phyto.df <- phyto.df %>%
  mutate( sampledate =  as.Date(sampledate)) %>%
  mutate(concat = paste0(station, sampledate))

#create concatinated id for water quality data
wq.df <- wq.df %>%
  mutate( sampledate =  as.Date(sampledate)) %>%
  mutate(concat = paste0(station, sampledate))

#create a vector to store concatinated ids from phyto events
concat_vector <- pull(phyto.df, concat)

#include only the data in water quality data that has a matching concat id with phyto events
wq.df <-wq.df %>%
  filter(concat %in% concat_vector) 
```

this snippet of Zach's code removes any row flagged in the problem column(anything other than na)
28 items are removed from the list
do we want this?
```{r}
# wq.df <- wq.df %>%
#   filter(is.na(problem)
        #) 


         #,
         #parameter %in% c("chla", "doc", "pheo", "salinity") ...removed because these are the only 4 parameters included so
         #the code doesn't actually do anything
         
#)
```

selects columns from wq (eventid recently added)
```{r}
wq.df <- wq.df %>%
  select(station, sampledate, layer, depth, parameter, measurevalue, eventid) %>% #eventid,
  mutate(sampledate=as.Date(sampledate))
```

```{r}
# step_3_test.df <- wq.df
# 
# data.table::fwrite(step_3_test.df, file.path(project.dir, "data/test", "step_3_test.csv"))
```


#### Prep Event

```{r}
phyto.df <- phyto.df %>%
  mutate(sampledate = as.Date(sampledate))
```

```{r}
event.df <- event.df %>%
  mutate(sampledate = as.Date(sampledate))
```

this is Zach's code for use with use for (I believe) older data that uses older salzone designations,
it's doing nothing for our current data set (2013-2016)
```{r}
event.df <- event.df %>%
  mutate(sampledate = as.Date(sampledate),
         salzone = case_when(
           salzone %in% c("tf", "fe") ~ "f",
           salzone %in% c("m", "me") ~ "m",
           salzone %in% c("o", "oe") ~ "o",
           salzone %in% c("p", "pe") ~ "p",
           TRUE ~ as.character(NA)
         ))
```

```{r}
pdepth.df <- event.df %>%
  filter(layer %in% c("ap","wc")) %>%
  mutate(sampledate = as.Date(sampledate)) %>%
  select(station, sampledate, pdepth) %>%
  dplyr::distinct()
```

COMPARE STEP 2
both 533
```{r}
step_2.df <- pdepth.df

data.table::fwrite(step_2.df, file.path(project.dir, "data/test", "step_2.csv"))

s2_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step2.csv")) %>% clean_up() #%>% mutate(sampledate=as.Date(sampledate))
```


applies pdepth to water quality dataframe
```{r}
wq_w_pdepth.df <- left_join(wq.df, pdepth.df, by = c("station", "sampledate"))
```

determine if depth is above pdepth
with exception for RET stations
```{r}
wq_w_pdepth.df <- wq_w_pdepth.df %>%
  mutate(above_pdepth = case_when(depth <= pdepth~"Yes",
                                  #depth > pdepth & station %in% c("ret3.1","ret4.3","ret3.1")~"No, but RET",
                                  depth > pdepth~"No"

                                  ))
```

```{r}
# test <- wq_w_pdepth.df %>%
#   filter(above_pdepth == "No, but RET")
```


COMPARE STEP 3
note: 7358 to 7358 in Claire's excel
```{r}
step_3.df <- wq_w_pdepth.df

data.table::fwrite(step_3.df, file.path(project.dir, "data/test", "step_3.csv"))

s3_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step3.csv")) %>% clean_up() #%>% mutate(sampledate=as.Date(sampledate))

```

test script for step 3
```{r}
# step_3_test.df <- step_3.df %>%
#   group_by(eventid) %>% #, sampledate) %>%
#   mutate(total_in_group = n()) %>%
#   summarise() %>%
#   ungroup()
#   
# s3_compare.df <- s3_compare.df %>%
#   mutate( sampledate =  as.Date(sampledate, format = '%m/%d/%Y'))
# 
# s3_compare_test.df <- s3_compare.df %>%
#   group_by(wq_eventid) %>%
#   mutate(total_in_group = n()) %>%
#   summarise() %>%
#   ungroup()
# 
# id_vector <- pull(s3_compare_test.df, wq_eventid)
# 
# step_3_by_id <-wq.df %>%
#   filter(eventid %in% id_vector)
# 
# step_3_to_exclude <- wq.df %>%
#   filter(!eventid %in% id_vector)
# 
# s3_wq_dates.df <- data.table::fread(file.path(project.dir, "data/test/", "wq_dates.csv")) %>% clean_up() %>%
#   mutate( sampledate =  as.Date(sampledate, format = '%m/%d/%Y'))
#   # group_by(sampledate) %>%
#   # mutate(total_in_group = n()) %>%
#   # summarise() %>%
#   # ungroup()
# 
# test_7358 <- step_3.df %>%
#   #filter(is.na(problem))
#   filter(layer %in% c("s","a","p","m")) %>%
# 
# test_missing <-  

  
```

filters out dpeths below pdepth
making an exception for any data missing pdepth of station tf4.2
by joining data that meet that criteria back in
```{r}
# missing_pdepth.df <- wq_w_pdepth.df %>%
#   filter( (is.na(pdepth)& station %in% c("tf3.3","tf4.2","tf5.5", "ret3.1","ret4.3","ret3.1")) 
# )
# 
# missing_other_pd.df <- left_join(missing_other.df, pdepth.df, by = c("station", "sampledate")) %>%
#   filter(totaldepth <= 5 & station %in% c("ret3.1","ret4.3") & pdepth <=1 ) 

```

```{r}
wq_above_pdepth.df <- wq_w_pdepth.df%>%
  filter(depth <= pdepth)


#wq_above_pdepth.df <- full_join(wq_above_pdepth.df, missing_pdepth.df)

#& (!station %in% c("tf3.3","tf4.2","tf5.5", "ret3.1","ret4.3","ret3.1") & layer != "ap"))


```


check for missing(station and date combos that are present in event.df but not represtented in wq_above_pdepth.df) 
```{r}
step_4.vec <- c("station", "sampledate", "layer", "pdepth", "concat", "depth", "parameter", "measurevalue", "above_pdepth", "eventid")

#function for finding backbone values missing from recently modified data and add them into recently modified(latest) date
find_missing <- function(data_backbone, source_data, modified_data, step_vector) {
  missing.df <- anti_join(data_backbone, modified_data, by = c("station","sampledate"))
  
  missing.df <-  missing.df %>%
    mutate( sampledate =  as.Date(sampledate)) %>%
    mutate(concat = paste0(station, sampledate))
  
  source_data <- source_data %>%
    mutate( sampledate =  as.Date(sampledate)) %>%
    mutate(concat = paste0(station, sampledate))
  
  concat_vector2.vec <- pull(missing.df, concat)
  
  missing_full.df <-source_data %>%
    filter(concat %in% concat_vector2.vec) 
  
  modified_data <- modified_data %>%
  mutate( sampledate =  as.Date(sampledate)) %>%
  mutate(concat = paste0(station, sampledate))

  final.df <- full_join(modified_data, missing_full.df, by = step_vector)
                          #c("station", "sampledate", "layer", "pdepth", "concat", "depth", "parameter", "measurevalue", "above_pdepth", "eventid"))
  
  return(final.df)
}
#show the mssing values
name_missing <- function(data_backbone, source_data, modified_data, step_vector) {
  missing.df <- anti_join(data_backbone, modified_data, by = c("station","sampledate"))
  
  missing.df <-  missing.df %>%
    mutate( sampledate =  as.Date(sampledate)) %>%
    mutate(concat = paste0(station, sampledate))
  
  source_data <- source_data %>%
    mutate( sampledate =  as.Date(sampledate)) %>%
    mutate(concat = paste0(station, sampledate))
  
  concat_vector2.vec <- pull(missing.df, concat)
  
  final.df <-source_data %>%
    filter(concat %in% concat_vector2.vec) 
  
    return(final.df)
}

wq_above_pdepth.df <- find_missing(event.df, wq_w_pdepth.df, wq_above_pdepth.df, step_4.vec)

# #finds instances of station/sampledate combo that are missing from wq
# missing.df <- anti_join(event.df,wq_above_pdepth.df, by = c("station", "sampledate"))
# 
# #standardize date format and add a unique identifier to missing 
# missing.df <- missing.df %>%
#   mutate( sampledate =  as.Date(sampledate)) %>%
#   mutate(concat = paste0(station, sampledate))
# 
# #standardize date format and add a unique identifier to wq
# wq_w_pdepth.df <- wq_w_pdepth.df %>%
#   mutate( sampledate =  as.Date(sampledate)) %>%
#   mutate(concat = paste0(station, sampledate))
# 
# #create a vector to store concatinated ids from the missing combos
# concat_vector2.vec <- pull(missing.df, concat)
# 
# #uses the combos in missing.df to retrieve the event data for the missing values
# missing_full.df <-wq_w_pdepth.df %>%
#   filter(concat %in% concat_vector2.vec) 
# 
# #adds a concat id to wq_above_pdepth
# wq_above_pdepth.df <- wq_above_pdepth.df %>%
#   mutate( sampledate =  as.Date(sampledate)) %>%
#   mutate(concat = paste0(station, sampledate))
# 
# #joins the missing values back into most recent wq df
# wq_above_pdepth.df <- full_join(wq_above_pdepth.df, missing_full.df, by = c("station", "sampledate", "layer", "pdepth", "concat", "depth", "parameter", "measurevalue", "above_pdepth", "eventid"))
```


COMPARE STEP 4
note: 3483 to 3483 in Claire's excel
```{r}
step_4.df <- wq_above_pdepth.df

data.table::fwrite(step_4.df, file.path(project.dir, "data/test", "step_4.csv"))

s4_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step4.csv")) %>% clean_up() #%>% mutate(sampledate=as.Date(sampledate))
```

code to compare dataframes for testing step 4:
```{r}
step_4_modified.df <- step_4.df %>%
  mutate(sampledate = as.Date(sampledate)) %>%
  #select(-sampledate, above_pdepth) %>%
  mutate(author = "Luke")

# s4_compare.df <- s4_compare.df %>%
#   mutate(sampledate = as.Date(sampledate))

s4_compare_modified.df <- s4_compare.df %>%
  rename(above_pdepth = `above_p_depth?`) %>%
  rename(eventid = wq_eventid) %>%
  rename(pdepth = p_depth) %>%
  select(station, sampledate, layer, depth, parameter, measurevalue, eventid, pdepth, above_pdepth) %>%
  #select(-sampledate, -above_pdepth) %>%
  mutate(author = "Claire")


```


the objects in diff_check_step_4.df were the 4 entries that exist in Claire's data but not in mine. Exceptions written for such cases, except one which does not exist in my data:
```{r}

#diff_check.df <- left_join(step_4_modified.df, s4_compare_modified.df)


LV_diff_check_step_4.df <- anti_join(step_4_modified.df, s4_compare_modified.df, by = c("station", "depth", "parameter", "measurevalue", "eventid", "pdepth")) #"layer",

C_diff_check_step_4.df <- anti_join(s4_compare_modified.df, step_4_modified.df,  by = c("station", "depth", "parameter", "measurevalue", "eventid", "pdepth")) #"layer",

# author_check.df <- diff_check.df %>%
#   filter(author == "Claire")
```


Step 5:

this code isolates exceptions that need to be added back in after the following chunk of code
```{r}
# 
# step5_exceptions <- wq_above_pdepth.df %>%
#   filter(above_pdepth == "No, but RET")
#   filter(!layer %in% c("s","f","m","p", "ap"))

```


generates new dataframe with new parameter s_chla (531 items, should be 533)
```{r}
s_chla.df <- wq_above_pdepth.df %>%
  filter(layer == "s", 
         parameter == "chla") %>% 
  unite(parameter, c("layer", "parameter"), remove = FALSE) #%>%
  # filter(depth <= pdepth) #%>%
  #filter(parameter=="s_chla")
  #filter(eventid==425680 | eventid==394855)
```

```{r}
# s_chla.df <- s_chla.df %>%
#   group_by(station, sampledate) %>%
#   mutate(avg_di_s_chla = mean(measurevalue, na.rm = TRUE)) %>%
#   ungroup()
```

Average replicates from s_chla ...measure value average
```{r}
di_s_chla.df <- s_chla.df %>%
  group_by(station, sampledate, depth) %>%
  #do not remove NAs
  mutate(measurevalue_averaged = mean(measurevalue)) %>%
  distinct(station, sampledate, depth, eventid, layer, parameter, measurevalue_averaged, concat) %>%
  ungroup()

```


find missing values in step 5
```{r}
step_5.vec <- c("station", "sampledate", "layer", "concat", "depth", "parameter", "eventid")


#grabs missing values
missing_s_chla.df <- name_missing(event.df, wq_above_pdepth.df, di_s_chla.df, step_5.vec) %>%
  #reduce to only a single isntance for each eventid
  distinct(eventid, station, sampledate, concat) %>%
  mutate(parameter =  "s_chla") %>%
  mutate(sampledate = as.Date(sampledate)) %>%
  mutate()

# creates a vector of the unqiue ids of the missing values
missing_ids.vec <- pull(missing_s_chla.df, eventid) 

# checks ids against source, which is totally redundant
missing_full.df <-wq_above_pdepth.df %>%
  filter(eventid %in% missing_ids.vec ) 

#join missing values into di_s_chla.df with blank values
di_s_chla.df <- full_join(di_s_chla.df, missing_s_chla.df, by = c("station", "sampledate", "eventid", "parameter", "concat"))

# test.df <- di_s_chla_test.df %>%
#   filter(is.na(measurevalue_averaged))
  
```


I have eventid 39855 with s_chla in my wq data
this is the check:
```{r}
mystery.df <- wq_above_pdepth.df %>%
  filter(eventid == 394855)

mystery2.df <- event.df %>%
  filter(station == "tf4.2" & sampledate =="2014-07-15")
  
mystery3.df <- wq.df %>%
  filter(eventid == 394855)

mystery4.df <- di_s_chla.df %>%
  filter(eventid == 394855)
```




STEP 5 COMPARE
my 533 to Claire's 531(was 533?)
```{r}
step_5.df <- di_s_chla.df

data.table::fwrite(step_5.df, file.path(project.dir, "data/test", "step_5.csv"))

s5_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step5.csv")) %>% clean_up() #%>% mutate(sampledate=as.Date(sampledate))
```

```{r}
step_5_modified.df <- step_5.df %>%
  mutate(sampledate = as.Date(sampledate)) %>%
  #select(-sampledate, above_pdepth) %>%
  mutate(author = "Luke") %>%
  clean_up() %>%
  mutate(measurevalue_averaged = as.numeric(measurevalue_averaged))

# s5_compare.df <- s5_compare.df %>%
#   mutate(sampledate = as.Date(sampledate))

s5_compare_modified.df <- s5_compare.df %>%
  rename(eventid = wq_eventid) %>%
  rename(measurevalue_averaged = average_of_measurevalue) %>%
  #rename(pdepth = p_depth) %>%
  #select(station, sampledate, layer, depth, parameter, measurevalue, eventid, pdepth, above_pdepth) %>%
  #select(-sampledate, -above_pdepth) %>%
  mutate(author = "Claire") %>%
  clean_up() %>%
  mutate(measurevalue_averaged = as.numeric(measurevalue_averaged))


C_diff_check_step_5.df <- anti_join(s5_compare_modified.df, step_5_modified.df, by = c("station",  "parameter", "measurevalue_averaged", "eventid", "layer"))

C_diff_check_step_5.df <- C_diff_check_step_5.df%>%
  mutate(measure_type = typeof(measurevalue_averaged)) %>%
  filter(eventid==425680 | eventid==394855)

LV_diff_check_step_5.df <- anti_join(step_5_modified.df, s5_compare_modified.df, by = c("station", "measurevalue_averaged", "parameter", "eventid","layer")) 

LV_diff_check_step_5.df <- LV_diff_check_step_5.df%>%
  mutate(measure_type = typeof(measurevalue_averaged))
```


```{r}
# wq_rep_check.df <- wq_above_pdepth.df %>%
#   select(station, sampledate, depth, parameter) %>%
#   distinct()
```



Averge replicates for the rest of the data ...measure value average 
```{r}
wq_above_pdepth.df <- wq_above_pdepth.df %>%
  group_by(station, sampledate, depth, parameter) %>%
  #do not remove NAs
  mutate(measurevalue_averaged = mean(measurevalue)) %>%
  distinct(station, sampledate, depth, parameter,measurevalue_averaged) %>%
  ungroup()
```

COMPARE STEP 6 
my 3366 to Claire's is 3357(was 3366)
```{r}
step_6.df <- wq_above_pdepth.df

data.table::fwrite(step_6.df, file.path(project.dir, "data/test", "step_6.csv"))

s6_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step6.csv")) %>% clean_up()

C_diff_check_step_6.df <- anti_join(step_6.df,s6_compare.df, by = c(
  #"station", "depth","parameter", 
  "measurevalue_averaged"
  )) 

LV_diff_check_step_6.df <- anti_join(s6_compare.df, step_6.df, by = c(
  #"station", "depth", "parameter", 
  "measurevalue_averaged")) 


```


```{r}
wq_above_pdepth.df <- wq_above_pdepth.df %>%
  group_by(station, sampledate, parameter) %>%
  mutate(depth_integrated_value = mean(measurevalue_averaged)) %>%
  distinct(station, sampledate, depth_integrated_value) %>%
  ungroup()
```

COMPARE STEP 7 
mine is 2028 to Claire's 2026(was 2028)
```{r}
step_7.df <- wq_above_pdepth.df

data.table::fwrite(step_7.df, file.path(project.dir, "data/test", "step_7.csv"))

s7_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step7.csv")) %>% clean_up() %>% 
  rename(depth_integrated_value =`depth-integrated_value`)


C_diff_check_step_7.df <- anti_join(step_7.df,s7_compare.df, by = c(#"station", "parameter", 
  "depth_integrated_value")) 

LV_diff_check_step_7.df <- anti_join(s7_compare.df, step_7.df, by = c(#"station", "parameter", 
  "depth_integrated_value")) 
```

join surface chloraphyl into water quality data
measurevalued_averaged is renamed depth_integrated_value to match water quality naming convention
(note:check against Claire's data to confirm that s_chla depth integrated values look right just to be sure)
```{r}
di_s_chla.df <- di_s_chla.df %>% 
  rename(depth_integrated_value = measurevalue_averaged)
  

wq_all.df <- full_join(wq_above_pdepth.df, di_s_chla.df, by = c("station", "sampledate", "parameter", "depth_integrated_value"))

# wq_all.df <- wq_all.df %>%
#   filter(parameter != "s_chla")
  
```


COMPARE STEP 8
my 2561 to Claire's 2555(was 2558)
```{r}
print(2028 + 533)


step_8.df <- wq_all.df

data.table::fwrite(step_8.df, file.path(project.dir, "data/test", "step_8.csv"))

s8_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step8.csv")) %>% clean_up() %>% 
  rename(depth_integrated_value =`depth-integrated_value`)

C_diff_check_step_8.df <- anti_join(step_8.df,s8_compare.df, by = c("depth_integrated_value")) 

LV_diff_check_step_8.df <- anti_join(s8_compare.df, step_8.df, by = c("depth_integrated_value")) 
```

filter out dates outside Mar-May and Jul-Sept
borrowed this chunk from Zach
```{r}
wq_warm.df <- wq_all.df %>%
  mutate(month = month(sampledate),
         season = case_when(
           month %in% c(3, 4, 5) ~ "spring",
           month %in% c(7, 8, 9) ~ "summer",
           TRUE ~ "remove"
          )) 
  #%>% 
  # filter(season %in% c("spring", "summer"))
```


assign salzones...work in progress
```{r}
wq_sal.df <- wq_warm.df %>%
  filter(parameter == "salinity")

wq_sal.df <- wq_sal.df %>%
  mutate(salzone = case_when(
    parameter == "salinity" & depth_integrated_value >0 & depth_integrated_value <= 0.5 ~ "F",
    parameter == "salinity" & depth_integrated_value > .5 &  depth_integrated_value <= 5 ~ "O",
    parameter == "salinity" & depth_integrated_value > 5 & depth_integrated_value <= 18 ~ "M",
    parameter == "salinity" & depth_integrated_value > 18~ "P")) %>%
    mutate(concat = paste0(station, sampledate)) 

wq_sal_id.df <- wq_sal.df %>%
  distinct(concat, salzone)

wq_warm.df <- wq_warm.df %>%
    mutate(concat = paste0(station, sampledate))   


wq_warm_sal.df <- left_join(wq_warm.df, wq_sal_id.df, by = c("concat"))#, "salzone"))



```




Find events without salinity and assign salinity when possible...Assign F salzone to some station/season instances based on known salinity profiles
```{r}
# salinities are less than 0.5 ppt, or "F" more than 90% of the time: CB1.1, ET5.1, TF1.5, TF2.3, TF4.2, TF5.5 in Spring;								
# CB1.1, TF2.3, TF4.2, TF5.5 in Summer. We can assume their Salzone is F and replace 61 of the 78 blanks.								
wq_no_sal.df <- anti_join(wq_warm_sal.df, wq_sal_id.df, by = c("concat"))

wq_sal_fill.df <-  wq_no_sal.df %>%
  distinct(concat, station, sampledate, season, salzone) %>%
  mutate(salzone = case_when((station %in% c("cb1.1", "tf2.3", "tf4.2","tf5.5" ) & is.na(salzone)) | (season == "spring" & station %in% c("et5.1", "tf1.5")& is.na(salzone))
         ~ "F", TRUE ~ salzone  ))

#test number of exceptions above code is catching
wq_sal_filled.df <- wq_sal_fill.df %>%
  filter(salzone == "F")

test_na.df <- wq_sal_fill.df %>%
  filter(is.na(salzone))
```

Add the data assigned salzones back into backbone
```{r}
wq_warm_sal_added.df <- left_join(wq_warm_sal.df, wq_sal_filled.df, by= c("concat", "salzone","station","sampledate", "season"))

# test to determine if we are retaining the right values
test.df <- anti_join(wq_warm.df, wq_warm_sal_added.df,  by = c("concat", "station","sampledate", "season", "parameter", "month", "depth_integrated_value"))
```



bring in Jackies data to compare salzones against
import jackie's salzones if data falls within the applicable date range and is missing salzone data
note:this should be the last method of acquiring salzone data implemented
(this code is untested)
```{r}
#jackie.df <- data.table::fread(file.path(project.dir, "data/Jackie/", "JMJ_PIBI_Salzone_Data.csv"))
jackie_salzones.df <- readxl::read_excel(file.path(project.dir, "data/Jackie/JMJ_PIBI_Salzone_Data.xlsx"),
                         sheet = "JMJ Salzone+Scores") %>% clean_up() %>%
  rename(salzone = ibi_salzone)

wq_need_sal.df <- wq_warm_sal_added.df %>%#wq_sal_test.df %>%
  mutate(year = year(sampledate)) %>%
  filter(year >= 1984 & year <= 2011 & is.na(salzone))

#join jackie salzone values to wq_need_sal by id, but only if sal needed is not an empty dataframe
if (nrow(wq_need_sal.df) > 1) {
  wq_has_sal.df <- left_join(wq_need_sal.df, jackie_salzones.df, by= c("concat", "salzone"))

  #join wq_need_sal to main wq data
  wq_warm_sal_added.df <- left_join(wq_warm_sal_added., wq_has_sal.df, by= c("concat", "salzone"))
}
             
```


```{r}
wq_warm.wide <- wq_warm_sal_added.df %>%
  select(-layer, -depth,-eventid,-concat) %>%
  spread(parameter, depth_integrated_value)
```


COMPARE STEP 9
```{r}
step_9.df <- wq_warm.wide

data.table::fwrite(step_9.df, file.path(project.dir, "data/test", "step_9.csv"))

s9_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step9.csv")) %>% clean_up() 

C_diff_check_step_9.df <- anti_join(step_9.df,s9_compare.df, by = c("chla", "s_chla", "pheo", "salinity", "doc")) 

LV_diff_check_step_9.df <- anti_join(s9_compare.df, step_9.df, by = c("chla", "s_chla", "pheo", "salinity", "doc")) 
```

remove winter and fall
```{r}
wq_final.df <- wq_warm.wide %>%

 filter(season %in% c("spring", "summer"))
```
COMPARE STEP 10

```{r}
step_10.df <- wq_final.df

data.table::fwrite(step_10.df, file.path(project.dir, "data/test", "step_10.csv"))

s10_compare.df <- data.table::fread(file.path(project.dir, "data/test/", "C_step10.csv")) %>% clean_up() 

C_diff_check_step_10.df <- anti_join(step_10.df,s10_compare.df, by = c("chla", "s_chla", "pheo", "salinity", "doc")) 

LV_diff_check_step_10.df <- anti_join(s10_compare.df, step_10.df, by = c("chla", "s_chla", "pheo", "salinity", "doc"))
```

itis
```{r}
# carbon.df <- readxl::read_excel(file.path(project.dir, "data/carbon/carbon_list_2014.xlsx"),
#                          sheet = "OldBiomassEstimates") %>% clean_up()

carbon.df <- data.table::fread(file.path(project.dir, "data/carbon/", "carbon_list_2014.csv")) %>% clean_up()
```

#### Prep Carbon Data

carbon data needs to be reformatted to work with bay data
```{r}
carbon.df <- carbon.df %>%
  mutate(tsn =  gsub("(?<![0-9])0+", "", tsn, perl = TRUE)) %>%
  rename(latinname = lbl) %>%
  rename(carbon = `carbon_(picogram_c_per_cell)`) %>%
  mutate(tsn = as.integer(tsn)) %>%
  mutate(size = case_when(size == "not_applicable"~ as.character(NA), TRUE ~size))

```

bring in phyto taxa data
```{r}
col.class.vec <- c("samplenumber" = "character",
                   "tsn" = "character",
                   "speccode" = "character")

taxa.raw <- data.table::fread(file.path(project.dir, "data/CEDR/cedr_phyto_taxa.csv"),
                            data.table = FALSE,
                            colClasses = col.class.vec,
                            na.strings = "") %>% 
  mutate(sampledate = as.Date(sampledate))
```

reduce to spring and summer and above pycnocline
```{r}
bay.df <- taxa.raw %>% 
  filter(layer %in% c("ap", "wc")) %>% 
  distinct() %>% 
  mutate(month = month(sampledate),
         season = case_when(
           month %in% c(3, 4, 5) ~ "spring",
           month %in% c(7, 8, 9) ~ "summer",
           TRUE ~ "remove"
         )) %>% 
  filter(season %in% c("spring", "summer")) %>%
  #changing types to match as is done in Zach's code using col.class.vec
  mutate(tsn=as.character(tsn)) %>%
  mutate(speccode = as.character(speccode)) %>%
  mutate(reportingvalue = as.integer(reportingvalue)) %>%
  mutate(samplenumber = as.character(samplenumber))

bay.df <- bay.df %>%
    mutate(size = case_when(size == "not_applicable"~ as.character(NA), TRUE ~size))
```

If there are multiple taxa with the same taxonomic name, then the reporting values should be summed. Each row should represent a unique taxon. (currently removes a single value)
```{r}
bay.df <- bay.df %>% 
  group_by_at(vars(-reportingvalue)) %>% 
  summarize(reportingvalue = sum(reportingvalue)) %>% 
  ungroup()
```

Only phytoplankton data (“ph”) is retained for analysis. Additionally, some taxa were not identified to any useful taxonomic rank and were excluded from the analysis.
```{r}
bay.df <- bay.df %>% 
  filter(grepl("ph", method),
         !latinname %in% c("micro-phytoflagellates",
                           "microflagellates",
                           #"green_cells",
                           #"blue_green_sphere",
                           "epiphytic_flagellates",
                           "hydrodictyon_reticulatum"))
```

```{r}
bay.df <- bay.df %>% 
  mutate(
    tsn = as.integer(tsn),
    tsn = case_when(
      latinname == "navicula_notablis" ~ as.integer(4327),
      latinname == "pleurosigma_macrum" ~ as.integer(4650),
      latinname == "pleurosigma_obscurum" ~ as.integer(591383),
      latinname == "polykrikos_hartmannii" ~ as.integer(331299),
      latinname == "protoperidinium_aciculiderum" ~ as.integer(10329),
      latinname == "protoperidinium_paulseni" ~ as.integer(3568),
      latinname == "scrippsiella_favionese" ~ as.integer(10537),
      latinname == "tetrastrum_caudatum" ~ as.integer(5691),
      latinname == "didymocystis" ~ as.integer(5810),
      latinname == "lauterborniella_elegantissima" ~ as.integer(6097),
      latinname == "characium_sp." ~ as.integer(5756),
      latinname == "cylindrospermopsis_sp." ~ as.integer(203689),
      latinname == "chaetoceros_neogracilis" ~ as.integer(1004011),
      latinname == "navicula_retusa_cancellata" ~ as.integer(1020372),
      latinname == "karlodinium_micrum" ~ as.integer(180904),
      latinname == "lagerheimia" ~ as.integer(6017),
      latinname == "quadricoccus_euryhalinicus" ~ as.integer(957939),
      latinname == "scrippsiella_precaria" ~ as.integer(10536),
      latinname == "psuedosolenia_calcar-avis" ~ as.integer(970064),
      latinname == "centronella" ~ as.integer(970064),
      latinname == "amphidinium_tatrae" ~ as.integer(9997),
      latinname == "navicula_lata" ~ as.integer(4450),
      latinname == "nitzschia_vitrea_recta" ~ as.integer(5204),
      latinname == "rhaphoneis_gemmifera" ~ as.integer(3145),
      latinname == "delphineis_surirella" ~ as.integer(969978),
      latinname == "navicula_annulata" ~ as.integer(3649),
      latinname == "proboscia_alata_gracillima" ~ as.integer(610099),
      latinname == "guinardia_striata" ~ as.integer(2921),
      latinname == "guinardia_cylindrus" ~ as.integer(2921),
      latinname == "aphanizomenon_issatschenkoi" ~ as.integer(1191),
      latinname == "helicotheca_tamesis" ~ as.integer(590815),
      latinname == "corethron_valdivae" ~ as.integer(2386),
      latinname == "gonyaulax_conjuncta" ~ as.integer(10359),
      latinname == "lioloma_delicatulum" ~ as.integer(573597),
      latinname == "syracosphaera_histrica" ~ as.integer(2234),
      latinname == "rhizosolenia_formosa" ~ as.integer(2879),
      latinname == "proboscla_alata_curvirostris" ~ as.integer(610099),
      latinname == "membraneis_challengeri" ~ as.integer(3648),
      latinname == "chrysococcus_tesselatus" ~ as.integer(1751),
      latinname == "rhoicosphenia_abbreviata" ~ as.integer(3633),
      latinname == "protoperidinium_aciculiferum" ~ as.integer(10340),
      latinname == "protoperidinium_fimbriatum" ~ as.integer(10340),
      latinname == "licmophora_inflata" ~ as.integer(3155),
      latinname == "biddulphia_reticulata" ~ as.integer(2678),
      latinname == "caloneis_lepidula" ~ as.integer(4369),
      latinname == "caloneis_trinodis" ~ as.integer(4369),
      latinname == "amphiprora_cholnokyi" ~ as.integer(4674),
      latinname == "navicula_interrupta" ~ as.integer(3649),
      latinname == "cerataulus_radiatus" ~ as.integer(2709),
      latinname == "gyrosigma_balticum_silimis" ~ as.integer(4623),
      latinname == "dictyocha_siderea" ~ as.integer(1804),
      latinname == "odontella_alternans" ~ as.integer(573604),
      latinname == "nitzschia_vitrea_salinarum" ~ as.integer(5204),
      latinname == "proboscla_alata_indica" ~ as.integer(610099),
      latinname == "attheya_decora" ~ as.integer(2876),
      latinname == "synedra_closterioides" ~ as.integer(970065),
      latinname == "trinacria_regina" ~ as.integer(2747),
      latinname == "chattonella" ~ as.integer(969917),
      latinname == "chattonella_subsalsa" ~ as.integer(969917),
      latinname == "heterosigma_akashiwo" ~ as.integer(969917),
      latinname == "vibrio_fisheri" ~ as.integer(959178),
      TRUE ~ as.integer(tsn)
    )
  )
```

carbon adjustments

L. this code requires the diameter variable from the carbon sheet in jackie's data.  We are currently using the biomass_estimates sheet.  To run this code(if it is determined to be necessary), you will need to import the carbon sheet from jackie's data and append the diameter variable through a join.
```{r}
# carbon.df <- carbon.df %>% 
#   mutate(size = case_when(
#     size %in% "unidentified" ~ as.character(NA),
#     str_detect(size, "species|sp.") ~ str_replace_all(size, "sp.#|species_#|species_|species", "sp#"),
#     TRUE ~size
#   ),
#   latinname = if_else(latinname == "blue_green_trichome", "blue_green_sphere", latinname)) %>% 
#   mutate(picoplankton = if_else(
#     diameter <= 20 &
#       height <= 20 &
#       width <= 20 &
#       width <= 20 &
#       height_2 <= 20 &
#       width_2 <= 20,
#     TRUE,
#     FALSE
#   )) %>% 
#   select(latinname, size, carbon) %>%
#   distinct() %>% 
#   bind_rows(data.frame(latinname = "asterionellopsis_glacialis", 
#                        carbon = carbon.df[carbon.df$latinname == "asterionellopsis_kariana", "carbon"],
#                        stringsAsFactors = FALSE))
```

check for duplicate values
```{r}
carbon.dups <- carbon.df %>% 
  group_by(latinname, size) %>% 
  mutate(count = n()) %>% 
  ungroup() %>% 
  filter(count > 1)
```

excluding 3 types that have dups
```{r}
carbon.df <- carbon.df %>% 
  filter(!(latinname == "chaetoceros_wighami" & is.na(carbon)),
         !(latinname == "biddulphia" & carbon == 7899.50),
         !(latinname == "gymnodinium" & carbon == 848))
```

Identify all taxa (latinname) in bay.df that do not have a size equivalent in carbon.df.

```{r}
no.match.df <- anti_join(bay.df, carbon.df, by = c("latinname", "size")) %>% 
  select(latinname, size) %>% 
  distinct()
```

```{r}
partial.match.df <- lapply(1:nrow(no.match.df), function(row.i) {
  sub.df <- no.match.df %>% 
    slice(row.i)
  
  size.vec <- str_split(sub.df$size, "_") %>% unlist()
  
  getmode <- function(v) {
    uniqv <- unique(v)
    uniqv[which.max(tabulate(match(v, uniqv)))]
  }
  
  
  carbon.sub <- carbon.df %>% 
    filter(latinname == sub.df$latinname) %>% 
    mutate(row_max_matches = sapply(size.vec, function(size.i) grep(size.i, size)) %>% 
             unlist() %>% 
             getmode()) %>% 
    slice(unique(row_max_matches)) %>% 
    select(-row_max_matches, -carbon) %>% 
    rename(new_size = size) %>% 
    mutate(size = sub.df$size)
  
  return(carbon.sub)
}) %>% 
  bind_rows()
```

```{r}
bay.df <- left_join(bay.df, partial.match.df, by = c("latinname", "size","tsn")) %>% 
  mutate(reported_size = size,
         size = if_else(!is.na(new_size), new_size, size))
```


note:not sure what,if anything, to do with this one:
Z. Some taxa require specific attention before they can be finalized.
```{r}
# bay.df <- bay.df %>%
#   mutate(
#     final_id = if_else(latinname == "trinacria_regina", "trinacria_regina",  final_id),
#     final_id = case_when(
#       latinname == "green_cells" ~ "green_cells",
#       latinname == "blue_green_spheres" ~ "blue_green_spheres",
#       TRUE ~ final_id
#     ),
#     kingdom = if_else(latinname == "trinacria_regina",
#                       "chromista", kingdom),
#     phylum = if_else(latinname == "trinacria_regina",
#                      "bacillariophyta", phylum),
#     subphylum = if_else(latinname == "trinacria_regina",
#                         "bacillariophytina", subphylum),
#     class = if_else(latinname == "trinacria_regina",
#                     "mediophyceae", class),
#     subclass = if_else(latinname == "trinacria_regina",
#                        "chaetocerotophycidae", subclass),
#     order = if_else(latinname == "trinacria_regina",
#                     "hemiaulales", order),
#     family = if_else(latinname == "trinacria_regina",
#                      "hemiaulaceae", family),
#     genus = case_when(latinname == "trinacria_regina" ~ "trinacria",
#                       latinname == "didymocystis" ~ "didymocystis",
#                       latinname == "Lauterborniella_elegantissima" ~ "lauterborniella",
#                       TRUE ~ genus),
#     species = case_when(latinname == "trinacria_regina" ~ "trinacria_regina",
#                         latinname == "lauterborniella_elegantissima" ~ "lauterborniella_elegantissima",
#                         TRUE ~ species)
#   )
```

fixes a spelling error in carbon_list (should fix in actual csv...)
other fixes to carbon_list such as assigning tsn to latinnames with known tsns
```{r}
carbon.df <- carbon.df %>%
  mutate(size = ifelse(size =="microphytoflaggelates", "microphytoflagellates", size)) %>%
  mutate(size = ifelse(size =="unidentified", as.character(NA), size)) %>%
  mutate(tsn = case_when(latinname == "aphanizomenon_issatschenkoi"~as.integer(1191), 
                         latinname == "chaetoceros_neogracilis"~as.integer(1004011), 
                         latinname =="guinardia_striata" ~ as.integer(2921), 
                         latinname == "lagerheimia" ~ as.integer(6017), 
                         latinname == "odontella_alternans"~ as.integer(573604),
                         #latinname == "protoperidinium_orbiculare" ~ as.integer(), 
                         latinname == "psuedosolenia_calcar-avis" ~ as.integer(970064),
                         
                         #over-writes tsn of 591037 for the itis tsn of 591383
                         latinname == "pleurosigma_obscurum" ~ as.integer(591383),
                         
                         TRUE ~ as.integer(tsn))) 
```

adds asterionellopsis_glacialis to the carbon.df with asterionellopsis_kariana carbon value
```{r}
carbon.df <- carbon.df %>%
  bind_rows(data.frame(latinname = "asterionellopsis_glacialis",
                       carbon = carbon.df[carbon.df$latinname == "asterionellopsis_kariana", "carbon"],
                       tsn = 615890,
                       group = "diatom",
                       stringsAsFactors = FALSE))
```

```{r}
# bay.df <- bay.df %>%
#   mutate(tsn = ifelse(tsn == 0, as.character(NA), tsn)) 

carbon.df <- carbon.df %>%
  mutate(tsn = ifelse(is.na(tsn)==TRUE, 0, tsn))

```

changes to bay data to match changes to carbon list
```{r}
bay.df <- bay.df %>%
  #mutate(group =ifelse(group =="chlorophycota", "chlorophyta", group )) %>%
  mutate(latinname = ifelse(latinname=="protoperidinium_bipes", "minuscula_bipes", latinname)) %>%
  mutate(latinname = ifelse(latinname=="woronichinia_naegeliana", "coelosphaerium_naegelianum", latinname)) %>%
  mutate(latinname = ifelse(latinname=="centritractis_belonophorus", "centratractus_belonophorus", latinname)) %>%
  mutate(tsn = ifelse(tsn==10210, 10209, tsn)) %>%
  mutate(tsn = ifelse(latinname=="woronichinia_naegeliana", 0, tsn)) %>%
  #modify centratractus_belonophorus tsn to match carbon_list
  mutate(tsn = ifelse(tsn==2014, 2013, tsn))
```


finally, join bay and carbon
```{r}
bay_merged.df <- left_join(bay.df, carbon.df, by = c("latinname","size","tsn")) %>%
  #remove dupliactes
  distinct()
```

we are generating duplicates on the above merge
```{r}
# bay_ex.df <- bay_merged.df %>%
#   distinct()
  #anti_join(bay.df, bay_merged.df,    by = c("station", "sampledate"))
```


clean up bay_merged(make group and nodc_phyla from carbon_list primary variable and remove value brought in from bay.df)
```{r}
bay_merged.df <- bay_merged.df %>%
  rename(group = group.y) %>%
  rename(nodc_phyla = nodc_phyla.y) %>%
  select(-group.x, -nodc_phyla.x)
```

turn carbon NA's into zeroes and calcualte biomass
```{r}
bay_merged.df <- bay_merged.df %>%
  mutate(carbon = ifelse(is.na(carbon)==TRUE, 0, carbon)) %>%
  mutate(nano_micro_biomass = reportingvalue * carbon / 10 ^ 6) 
```

merge wq and main data
```{r}
bay_merged.df <-  bay_merged.df %>%
  mutate(concat = paste0(station, sampledate)) 

wq_final.df <- wq_final.df %>%
  mutate(concat = paste0(station, sampledate)) 

bay_wq_merged.df <- left_join(bay_merged.df, wq_final.df, by = c("concat", "station","sampledate", "month", "season"))
```

```{r}
#tf3.3_ap_fs1_2016-03-17
#2013-03-05

test <-  bay_merged.df %>%
  filter(station == "tf3.3" & sampledate == "2016-03-17")
```


#### Calulate Metrics
```{r}
bay_wq_merged_calc.df <-  bay_wq_merged.df %>%
  
  #group by event
  group_by(concat) %>%
  
  
  #number of members in group
  # mutate(count = n()) %>%
  #name of new biomass?
  mutate(total_biomass = sum(carbon)) %>%
  mutate(total_nano_micro_biomass = sum(nano_micro_biomass)) %>%
  #
  
  #mutate(carbon_chlorophyll_ratio_wrong = nano_micro_biomass/chla) %>% #invert division to get reverse
  mutate(carbon_chlorophyll_ratio = total_nano_micro_biomass/chla) %>%
  
  group_by(group, add=TRUE) %>%

  mutate(diatom_biomass = case_when(group == "diatom"~sum(carbon)/1000000, TRUE ~ as.numeric(NA))) %>%
  mutate(dinoflagellate_biomass = case_when(group == "dinoflagellates"~sum(carbon)/1000000, TRUE ~ as.numeric(NA))) %>%
  mutate(cyanophyte_biomass = case_when(group == "cyanophyte"| group == "cyanobacteria"~sum(carbon)/1000000, TRUE ~ as.numeric(NA))) %>%
  mutate(cryptophyte_biomass = case_when(group == "cryptophyte" | group == "cryptophyceae" | group == "cryptophyta" | group == "cryptomonads"~sum(carbon), TRUE ~ as.numeric(NA))) %>%
  
  mutate(percent_cryptophyte_biomass = cryptophyte_biomass/total_nano_micro_biomass)%>%
  
  # #how should this be calculated?
  # mutate(new_metric = other_biomass*micro_biomass/count) %>%
  ungroup()


# test <- bay_wq_merged_calc.df %>%
#   filter(group == "cyanoophytes")
```


```{r}
test <- bay_wq_merged_calc.df %>%
  filter((group == "diatom" & concat == "tf5.52013-03-05") |  (group == "diatom" & concat =="sbe52013-03-13"))
```

```{r}
test <- bay.df %>%
  filter(latinname == "microcystis_aeruginosa")
```


```{r}
# groups <- bay_wq_merged_calc.df %>%
#   select(group) %>%
#   distinct()
```

```{r}
bay_wq_merged_calc.df <- bay_wq_merged_calc.df %>%
  group_by(concat) %>%
  mutate(prorocentrum_minimum_abundance = case_when(latinname == "prorocentrum_minimum"~ sum(as.double(reportingvalue)), TRUE ~ as.numeric(NA))) %>%
  mutate(microcystis_aeruginosa_abundance = case_when(latinname == "microcystis_aeruginosa"~ sum(as.double(reportingvalue)), TRUE ~ as.numeric(NA))) %>%
  ungroup()
```

```{r}
# bay_test <- bay_wq_merged_calc.df %>%
#   filter(latinname == "microcystis_aeruginosa" | latinname == "prorocentrum_minimum")  
```

join the picoplankton data into the bay data and average duplicates and generate picoplankton_abundance
```{r}

picoplankton.df <- data.table::fread(file.path(project.dir, "data/CEDR/", "cedr_pico_taxa.csv")) %>% clean_up() %>%
  rename(pico_reportingvalue = reportingvalue) %>%
  select("station", "sampledate", "pico_reportingvalue") %>% 
  mutate(sampledate=as.Date(sampledate)) %>%
  mutate(concat = paste0(station, sampledate)) %>%
  select(-"station",-"sampledate") %>%
  # take the average of reportingvalue where a unique station date combo has more than one reporting value
  group_by(concat) %>%
  mutate(pico_reportingvalue = mean(pico_reportingvalue)) %>%
  mutate(picoplankton_abundance = sum(pico_reportingvalue)) %>%
  unique() %>%
  ungroup()

bay_full_calc.df <- left_join(bay_wq_merged_calc.df, picoplankton.df,   by = c("concat"))



```

#### Scoring

```{r}
  metric.vec <- c("total_phyto_biomass_chla_ratio",
                  "surface_chla",
                  "chla",
                  "doc",
                  "pheophytin",
                  "total_phyto_biomass", 
                  "diatom_biomass",
                  "dinoflagellate_biomass",
                  "cyanophyte_biomass",
                  "cryptophyte_biomass",
                  "pct_cryptophyte",
                  "prorocentrum_minimum_abundance", 
                  "microcystis_aeruginosa_abundance",
                  #"pico_reportingvalue" 
                  "picoplankton_abundance"
                  
                  )
```

add a unique_id to match Zach's scoring function later on
replacing concat with it
```{r}
bay_full_calc.df <- bay_full_calc.df %>%
  mutate(unique_id = paste(station, layer, samplenumber, sampledate, sep = "_")) %>%
  select(-concat)
```

rename some metrics to fit Zach's code(that I have rewritten now)
```{r}
metrics_prep.df <- bay_full_calc.df %>%
  rename(pheophytin = pheo) %>% 
  rename(total_phyto_biomass = total_nano_micro_biomass) %>%
  rename(total_phyto_biomass_chla_ratio = carbon_chlorophyll_ratio) %>%
  rename(pct_cryptophyte = percent_cryptophyte_biomass) %>%
  rename(surface_chla = s_chla) %>%
  mutate(salzone = tolower(salzone)) 
  
```

```{r}
test <- metrics_prep.df %>%
  select(salzone)
```


```{r}
test <- metrics_prep.df %>%
  filter( (season == "summer" & salzone == "F") | (season == "summer" & salzone == "f")) %>%
  distinct()
```

try selecting only relevant fields
```{r}
metrics_prep.df <- metrics_prep.df %>%
  select(unique_id, station, sampledate, samplenumber, season, salzone, total_phyto_biomass_chla_ratio,
                  surface_chla,
                  chla,
                  doc,
                  pheophytin,
                  total_phyto_biomass, 
                  diatom_biomass,
                  dinoflagellate_biomass,
                  cyanophyte_biomass,
                  cryptophyte_biomass,
                  pct_cryptophyte,
                  prorocentrum_minimum_abundance, 
                  microcystis_aeruginosa_abundance,
                  picoplankton_abundance)
```


```{r}

score_spring_f <- metrics_prep.df %>% 
  filter(
           season == "spring",
           salzone == "f"#,

           # is.na(total_phyto_biomass_chla_ratio) == FALSE #|
           # is.na(surface_chla) == FALSE |
           # is.na(cyanophyte_biomass) == FALSE |
           # is.na(doc) == FALSE |
           # is.na(pheophytin) == FALSE |
           # is.na(total_phyto_biomass) == FALSE
           ) %>%
  mutate(total_phyto_biomass_chla_ratio_rating = case_when(
      total_phyto_biomass_chla_ratio < 39.28 ~1,
      total_phyto_biomass_chla_ratio>= 39.28 & total_phyto_biomass_chla_ratio <= 41.72 ~ 3,
      total_phyto_biomass_chla_ratio > 41.72 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(surface_chla_rating = case_when(
      surface_chla < 3.42 & surface_chla > 14.45~1,
      (surface_chla >= 3.42 & surface_chla <= 4.37) |(surface_chla >= 13.98 & surface_chla <= 14.45) ~ 3,
      surface_chla > 4.37 & surface_chla < 13.98 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(cyanophyte_biomass_rating = case_when(
      cyanophyte_biomass > 23.02 ~ 1,
      cyanophyte_biomass <= 23.02 ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 2.40 ~ 1,
      doc >= 2.19 & doc <= 2.40 ~ 3,
      doc  < 2.19 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 2.50 ~ 1,
      pheophytin >= 1.55 & pheophytin <= 2.50 ~ 3,
      pheophytin < 1.55 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass <= 172.99 | total_phyto_biomass > 828.5 ~ 1,
      total_phyto_biomass >= 583.9 & total_phyto_biomass <= 828.5 ~ 3,
      total_phyto_biomass > 172.99 & total_phyto_biomass < 583.9 ~ 5,
      TRUE ~ as.numeric(NA))) 
  
  

```

```{r}
score_spring_o <- metrics_prep.df %>% 
  filter(
           season == "spring",
           salzone == "o"#,

           ) %>%
  mutate(total_phyto_biomass_chla_ratio_rating = case_when(
      total_phyto_biomass_chla_ratio < 18.8 | total_phyto_biomass_chla_ratio > 48.6 ~ 1,
      total_phyto_biomass_chla_ratio >= 18.8 & total_phyto_biomass_chla_ratio <= 21.2 ~ 3,
      total_phyto_biomass_chla_ratio > 21.2 & total_phyto_biomass_chla_ratio < 48.6 ~ 5,
      TRUE ~ as.numeric(NA))) %>%

  mutate(surface_chla_rating = case_when(
      surface_chla < 6.77 | surface_chla > 33.64 ~ 1,
      (surface_chla >= 20.93 & surface_chla <= 33.64) |(surface_chla >= 6.77 & surface_chla <= 8.82) ~ 3,
      surface_chla > 8.82 & surface_chla < 20.93 ~ 5,
      TRUE ~ as.numeric(NA))) %>%

  mutate(doc_rating = case_when(
      doc > 3.27 ~ 1,
      doc >= 2.69 & doc <= 3.27 ~ 3,
      doc < 2.69 ~ 5,
      TRUE ~ as.numeric(NA))) %>%

  mutate(pheophytin_rating = case_when(
      pheophytin > 2.68 ~ 1,
      pheophytin >= 2.23 & pheophytin <= 2.68 ~ 3,
      pheophytin < 2.23 ~ 5,
    TRUE ~ as.numeric(NA))) %>%
  
  mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass > 133.67 & total_phyto_biomass < 426.31 ~ 5,
      (total_phyto_biomass >= 131.37 & total_phyto_biomass <= 133.67) | (total_phyto_biomass >= 426.31 & total_phyto_biomass < 685.81) ~ 3,
      total_phyto_biomass < 131.37 | total_phyto_biomass >= 685.81 ~ 1,
      TRUE ~ as.numeric(NA))) 
```
  
```{r}
score_spring_m <- metrics_prep.df %>% 
  filter(
           season == "spring",
           salzone == "m"#,

           ) %>%
  mutate(total_phyto_biomass_chla_ratio_rating = case_when(
       total_phyto_biomass_chla_ratio < 45.04 ~ 1,
      total_phyto_biomass_chla_ratio >= 45.04 & total_phyto_biomass_chla_ratio <= 69.52 ~ 3,
      total_phyto_biomass_chla_ratio > 69.52 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(surface_chla_rating = case_when(
      surface_chla < 2.60 | surface_chla > 8.00 ~ 1,
      (surface_chla >= 2.60 & surface_chla <= 2.90) |(surface_chla >= 6.17 & surface_chla <= 8.00) ~ 3,
      surface_chla > 2.90 & surface_chla < 6.17 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  
  mutate(diatom_biomass_rating = case_when(
      diatom_biomass < 149 | diatom_biomass >= 2513 ~ 1,
      diatom_biomass >= 149 & diatom_biomass <= 275.4 ~ 3,
      diatom_biomass > 275.4 & diatom_biomass < 2513 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(dinoflagellate_biomass_rating = case_when(
      dinoflagellate_biomass < 28.3 | dinoflagellate_biomass > 268.2 ~ 1,
      dinoflagellate_biomass >= 156.9 & dinoflagellate_biomass <= 268.2 ~ 3,
      dinoflagellate_biomass > 28.3 & dinoflagellate_biomass < 156.9 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 3.17 ~ 1,
      doc >= 2.84 & doc <= 3.17 ~ 3,
      doc < 2.84 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 1.03 ~ 1,
      pheophytin >= 1.00 & pheophytin <= 1.03 ~ 3,
      pheophytin < 1.00 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(prorocentrum_minimum_abundance_rating = case_when(
      prorocentrum_minimum_abundance > 1477600 ~ 1,
      prorocentrum_minimum_abundance <= 1477600 ~ as.numeric(NA),
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass > 1150 ~ 1,
      total_phyto_biomass <= 1150 ~ as.numeric(NA),
      TRUE ~ as.numeric(NA))) #%>%
```

```{r}
score_spring_p <- metrics_prep.df %>% 
  filter(
           season == "spring",
           salzone == "p"#,

           ) %>%
  mutate(total_phyto_biomass_chla_ratio_rating = case_when(
      total_phyto_biomass_chla_ratio < 71.0 ~ 1,
      total_phyto_biomass_chla_ratio >= 71.0 & total_phyto_biomass_chla_ratio <= 107.5 ~ 3,
      total_phyto_biomass_chla_ratio > 107.5 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(surface_chla_rating = case_when(
      surface_chla > 4.00 ~ 1,
      surface_chla >= 2.80 & surface_chla <= 4.00 ~ 3,
      surface_chla < 2.80 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pct_cryptophyte_rating = case_when(
      pct_cryptophyte > 7.06 ~ 1,
      pct_cryptophyte >= 4.93 & pct_cryptophyte <= 7.06 ~ 3,
      pct_cryptophyte < 4.93 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 2.61 ~ 1,
      doc >= 2.50 & doc <= 2.61 ~ 3,
      doc < 2.50 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 0.90 ~ 1,
      pheophytin >= 0.55 & pheophytin <= 0.90 ~ 3,
      pheophytin < 0.55 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(prorocentrum_minimum_abundance_rating = case_when(
      prorocentrum_minimum_abundance > 7488 ~ 1,
      prorocentrum_minimum_abundance >= 672 & prorocentrum_minimum_abundance <= 7488 ~ 3,
      prorocentrum_minimum_abundance < 672 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass > 1061.7 ~ 1,
      total_phyto_biomass <= 1061.7 ~ as.numeric(NA),
      TRUE ~ as.numeric(NA))) 
  
```

```{r}
score_summer_f <- metrics_prep.df %>% 
  filter(
           season == "summer",
           salzone == "f"#,

           ) %>%
  
  mutate(surface_chla_rating = case_when(
      surface_chla > 12.30 ~ 1,
      (surface_chla >= 12.00 & surface_chla <= 12.30) | surface_chla <= 5.4 ~ 3,
      surface_chla > 5.40 & surface_chla < 12.00 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(cyanophyte_biomass_rating = case_when(
      cyanophyte_biomass > 67.4 ~ 1,
      cyanophyte_biomass >= 38.87 & cyanophyte_biomass <= 67.4 ~ 3,
      cyanophyte_biomass < 38.87 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(diatom_biomass_rating = case_when(
      diatom_biomass > 192.6 ~ 1,
      diatom_biomass >= 122.1 & diatom_biomass <= 192.6 ~ 3,
      diatom_biomass < 122.1 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 3.18 ~ 1,
      doc >= 2.67 & doc <= 3.18 ~ 3,
      doc < 2.67 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(microcystis_aeruginosa_abundance_rating = case_when(
      microcystis_aeruginosa_abundance > 262507 ~ 1,
      microcystis_aeruginosa_abundance <= 262507 ~ as.numeric(NA),
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 4.30 ~ 1,
      pheophytin >= 2.40 & pheophytin <= 4.30 ~ 3,
      pheophytin < 2.40 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass > 555.7 ~ 1,
      total_phyto_biomass < 231.3 ~ 3,
      total_phyto_biomass >= 231.3 & total_phyto_biomass <= 555.7 ~ 5,
      TRUE ~ as.numeric(NA))) 
```

```{r}
score_summer_o <- metrics_prep.df %>% 
  filter(
           season == "summer",
           salzone == "o"#,

           ) %>%
  mutate(surface_chla_rating = case_when(
      surface_chla >= 9.47 ~ 1,
      surface_chla <= 4.20 ~ 3,
      surface_chla > 4.20 & surface_chla < 9.47 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(cyanophyte_biomass_rating = case_when(
      cyanophyte_biomass >= 26.55 ~ 1,
      cyanophyte_biomass < 1.79 ~ 3,
      cyanophyte_biomass >= 1.79 & cyanophyte_biomass < 26.55 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(diatom_biomass_rating = case_when(
      diatom_biomass >= 126.59 ~ 1,
      diatom_biomass <= 44.14 ~ 3,
      diatom_biomass > 44.14 & diatom_biomass < 126.59 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 4.00 ~ 1,
      doc >= 3.15 & doc <= 4.00 ~ 3,
      doc < 3.15 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 2.81 ~ 1,
      pheophytin >= 1.58 & pheophytin <= 2.81 ~ 3,
      pheophytin < 1.58 ~ 5,
      TRUE ~ as.numeric(NA))) 
  
```

```{r}
score_summer_m <- metrics_prep.df %>% 
  filter(
           season == "summer",
           salzone == "m"#,

           ) %>%
  mutate(total_phyto_biomass_chla_ratio_rating = case_when(
      total_phyto_biomass_chla_ratio < 32.2 ~ 1,
      total_phyto_biomass_chla_ratio >= 32.2 & total_phyto_biomass_chla_ratio <= 36.9 ~ 3,
      total_phyto_biomass_chla_ratio > 36.9 ~ 5,
      TRUE ~ as.numeric(NA))) %>%

  mutate(surface_chla_rating = case_when(
      surface_chla >= 9.74 ~ 1,
      (surface_chla >= 7.70 & surface_chla < 9.74) | surface_chla <= 4.00 ~ 3,
      surface_chla > 4.00 & surface_chla < 7.70 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(dinoflagellate_biomass_rating = case_when(
      dinoflagellate_biomass <= 31.22 | dinoflagellate_biomass > 200.92 ~ 1,
      dinoflagellate_biomass > 31.22 & dinoflagellate_biomass<= 55.98 ~ 3, 
      dinoflagellate_biomass > 55.98 & dinoflagellate_biomass < 200.92 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 3.35 ~ 1,
      doc >= 2.99 & doc <= 3.35 ~ 3,
      doc < 2.99 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 1.60 ~ 1,
      pheophytin >= 1.23 & pheophytin <= 1.60 ~ 3,
      pheophytin < 1.23 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(picoplankton_abundance_rating = case_when(
      picoplankton_abundance < 352000000 ~ 1,
      picoplankton_abundance >= 352000000 & picoplankton_abundance <= 598720000 ~ 3,
      picoplankton_abundance > 598720000 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass > 660 ~ 1,
      total_phyto_biomass <= 660 ~ as.numeric(NA),
      TRUE ~ as.numeric(NA))) 
  

```

```{r}
score_summer_p <- metrics_prep.df %>% 
  filter(
           season == "summer",
           salzone == "p"#,

           ) %>%
  mutate(total_phyto_biomass_chla_ratio_rating = case_when(
      total_phyto_biomass_chla_ratio < 37.7 ~ 1,
      total_phyto_biomass_chla_ratio >= 37.7 & total_phyto_biomass_chla_ratio <= 74.5 ~ 3,
      total_phyto_biomass_chla_ratio > 74.5 ~ 5,
      TRUE ~ as.numeric(NA))) %>%

  mutate(surface_chla_rating = case_when(
      surface_chla > 5.33 ~ 1,
      surface_chla >= 4.52 & surface_chla <= 5.33 ~ 3,
      surface_chla < 4.52 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pct_cryptophyte_rating = case_when(
      pct_cryptophyte > 6.5 ~ 1,
      pct_cryptophyte >= 3.9 & pct_cryptophyte <= 6.5 ~ 3,
      pct_cryptophyte< 3.9 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(diatom_biomass_rating = case_when(
      diatom_biomass >= 799 | diatom_biomass < 137 ~ 1,
      diatom_biomass >= 137 & diatom_biomass <= 181 ~ 3,
      diatom_biomass > 181 & diatom_biomass < 799 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(dinoflagellate_biomass_rating = case_when(
      dinoflagellate_biomass < 23 | dinoflagellate_biomass >= 544 ~ 1,
      dinoflagellate_biomass >= 23 & dinoflagellate_biomass <= 37 ~ 3, 
      dinoflagellate_biomass > 37 & dinoflagellate_biomass < 554 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(doc_rating = case_when(
      doc > 2.80 ~ 1,
      doc >= 2.58 & doc <= 2.80 ~ 3,
      doc < 2.58 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(pheophytin_rating = case_when(
      pheophytin > 1.50 ~ 1,
      pheophytin >= 0.93 & pheophytin <= 1.50 ~ 3,
      pheophytin < 0.93 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
  mutate(picoplankton_abundance_rating = case_when(
      picoplankton_abundance < 208600000 ~ 1,
      picoplankton_abundance >= 208600000 & picoplankton_abundance <= 269500000 ~ 3,
      picoplankton_abundance > 269500000 ~ 5,
      TRUE ~ as.numeric(NA))) %>%
  
    mutate(total_phyto_biomass_rating = case_when(
      total_phyto_biomass < 181 | total_phyto_biomass > 831 ~ 1,
      (total_phyto_biomass >= 181 & total_phyto_biomass <= 207) | (total_phyto_biomass >= 718 & total_phyto_biomass <= 831) ~ 3,
      total_phyto_biomass < 718 ~ 5,
      TRUE ~ as.numeric(NA))) 
```

```{r}
all_seasons <- bind_rows(score_spring_f, score_spring_o ,score_spring_m, score_spring_p, score_summer_f, score_summer_o, score_summer_m, score_summer_p)
```

```{r}
test <-  bay_full_calc.df %>%
  filter(is.na(season) | is.na(salzone))
```


```{r}
#disable scientific notation
options(scipen=999)




test_all_seasons <- all_seasons 

#this is not what we want to do
test_all_seasons[is.na(test_all_seasons)] <- 0 

scores_ratings_final <- test_all_seasons%>%
    # mutate(ibi_score =  case_when(season == "spring" & salzone == "f" ~ mean(total_phyto_biomass_chla_ratio_rating, surface_chla_rating, cyanophyte_biomass_rating, doc_rating, pheophytin_rating, total_phyto_biomass_rating, na.rm = T), TRUE~ as.numeric(NA)))
  
  mutate(ibi_score =  case_when(season == "spring" & salzone == "f" ~ (total_phyto_biomass_chla_ratio_rating + surface_chla_rating + cyanophyte_biomass_rating + doc_rating + pheophytin_rating + total_phyto_biomass_rating) / 6,
                                
                                  season == "spring" & salzone == "o" ~ (total_phyto_biomass_chla_ratio_rating + surface_chla_rating + doc_rating + pheophytin_rating +  total_phyto_biomass_rating) / 5,
                                
                                  season == "spring" & salzone == "m" ~ (total_phyto_biomass_chla_ratio_rating + surface_chla_rating + diatom_biomass_rating + dinoflagellate_biomass_rating + doc_rating + pheophytin_rating + prorocentrum_minimum_abundance_rating + total_phyto_biomass_rating) / 8,
                                
                                  season == "spring" & salzone == "p" ~ (total_phyto_biomass_chla_ratio_rating + surface_chla_rating + pct_cryptophyte_rating + doc_rating + pheophytin_rating + prorocentrum_minimum_abundance_rating + total_phyto_biomass_rating) / 7,
                                
                                
                                  season == "summer" & salzone == "f" ~ ( surface_chla_rating + cyanophyte_biomass_rating + diatom_biomass_rating + doc_rating + microcystis_aeruginosa_abundance_rating + pheophytin_rating + total_phyto_biomass_rating) / 7,
                                
                                  season == "summer" & salzone == "o" ~ (surface_chla_rating + cyanophyte_biomass_rating + diatom_biomass_rating + doc_rating + pheophytin_rating) / 5,
                                
                                  season == "summer" & salzone == "m" ~ (total_phyto_biomass_chla_ratio_rating + surface_chla_rating + dinoflagellate_biomass_rating + doc_rating + pheophytin_rating + picoplankton_abundance_rating + total_phyto_biomass_rating) / 7,
                                
                                  season == "summer" & salzone == "p" ~ (total_phyto_biomass_chla_ratio_rating +
surface_chla_rating + pct_cryptophyte_rating + diatom_biomass_rating + dinoflagellate_biomass_rating + doc_rating + pheophytin_rating + picoplankton_abundance_rating + total_phyto_biomass_rating) / 9,
                                
                                
                                
                                                TRUE~ as.numeric(NA)))
```

```{r}

scores_ratings_final %>% 
data.table::fwrite(file.path(rprojroot::find_rstudio_root_file(), "data/ScoresRatings", "final_scores_ratings.csv"))
```


